Simulating High-Dimensional Multivariate Data

### 2020-12-25 030-bigsimr-package.Rmd
4th draft. Address Ed comments Part I.
double spacing and updated bigsimr code

*Checking the theoretical correlation bounds*.
As discussed in Section \@ref(background), given a pair of marginal distributions the possible correlations are not free to vary between $[-1, 1]$. 
`bigsimr::cor_bounds` estimates the theoretic lower and upper correlation bounds, using the Generate, Sort, and Correlate algorithm of @DH2011.

```{r ch030-cor-bounds, echo=TRUE}
cor_bounds(margins = margins, type = type)
```

@@

```{r ch030-cor-bounds-pearson, echo=TRUE}
cor_bounds(margins = margins, type = 'pearson', reps = 1e6)
```

@@

```{r ch030-sim-margins, echo=TRUE}
x <- rvec(10000, rho, margins, type)
df_sim <- as.data.frame(x)
colnames(df_sim) <- colnames(df)
```

>>>>

*Checking the theoretical correlation bounds*.
As discussed in Section \@ref(background), given a pair of marginal distributions the possible correlations are not free to vary between $[-1, 1]$. 
`bigsimr::cor_bounds` estimates the pairwise lower and upper correlation bounds, using the Generate, Sort, and Correlate algorithm of @DH2011.

```{r ch030-cor-bounds, echo=TRUE}
cor_bounds(margins = margins, method = type)
```

@@

```{r ch030-cor-bounds-pearson, echo=TRUE}
cor_bounds(margins = margins, method = 'pearson', reps = 1e6)
```

@@

```{r ch030-sim-margins, echo=TRUE}
x <- rvec(10000, rho, margins)
df_sim <- as.data.frame(x)
colnames(df_sim) <- colnames(df)
```

### 2020-12-25 050-examples.Rmd
double spacing and updated bigsimr code

```{r ch050-quadloss, echo=F, eval=T, cache=FALSE}
quadLossRDS <- paste0( "results/brca", round(myProb*100), "quadLoss.rds" )
if ( !file.exists( quadLossRDS) )  {
    n <- nrow(brca)
    ## convert outside for faster simulation
    nb_Rho_p <- bigsimr::cor_convert( rho = nb_Rho,
                                     from = corType, to = "pearson" )
    ## ensure PSD
    nb_Rho_p <- bigsimr::cor_nearPSD( G = nb_Rho_p )
    ## create m random vectors and estimate correlation
    simRho <- replicate(n = m,
        expr = { tmpSim <- rvec(n = n , nb_Rho_p, nb_margins,
          type = 'pearson', ensure_PSD = FALSE, cores = cores);
          bigsimr::cor_fast( x = tmpSim, method = corType )} ,
     simplify = FALSE)
    ## find quadratic loss at each rep
    quadLoss <- unlist( lapply( simRho, rags2ridges::loss,
                               T = nb_Rho, type = "quadratic"))
    saveRDS(quadLoss, quadLossRDS)
}
quadLoss <- readRDS( quadLossRDS )
```

>>>>

```{r ch050-quadloss, echo=F, eval=T, cache=FALSE}
quadLossRDS <- paste0( "results/brca", round(myProb*100), "quadLoss.rds" )
if ( !file.exists( quadLossRDS) )  {
    n <- nrow(brca)
    ## convert outside for faster simulation
    nb_Rho_p <- bigsimr::cor_convert( rho = nb_Rho,
                                     from = corType, to = "pearson" )
    ## ensure PSD
    nb_Rho_p <- bigsimr::cor_nearPSD( G = nb_Rho_p )
    ## create m random vectors and estimate correlation
    simRho <- replicate(n = m,
        expr = { tmpSim <- rvec(n = n , nb_Rho_p, nb_margins,
          ensure_PSD = FALSE, cores = cores);
          bigsimr::cor_fast( x = tmpSim, method = corType )} ,
     simplify = FALSE)
    ## find quadratic loss at each rep
    quadLoss <- unlist( lapply( simRho, rags2ridges::loss,
                               T = nb_Rho, type = "quadratic"))
    saveRDS(quadLoss, quadLossRDS)
}
quadLoss <- readRDS( quadLossRDS )
```

### 2020-12-25 index.Rmd
double spacing and updated bigsimr code

--- 
title: "Simulating High-Dimensional Multivariate Data"
subtitle: "using the bigsimr R Package"
author:
  - Alfred G. Schissler
  - Edward J. Bedrick
  - Alexander D. Knudson
  - Tomasz J. Kozubowski
  - Tin Nguyen
  - Anna K. Panorska
  - Juli Petereit
  - Walter W. Piegorsch
  - Duc Tran
bibliography: ["bigsimr.bib", "packages.bib"]
abstract:  It is critical to realistically simulate data when conducting Monte Carlo studies and evaluating quantitative methods. But measurements are often correlated and high dimensional in this era of big data, however, such as data obtained through high-throughput biomedical experiments. Due to computational complexity and a lack of user-friendly software available to simulate these massive multivariate constructions, researchers may resort to simulation designs that posit independence or perform arbitrary data transformations. This article introduces the `bigsimr` R package that provides a flexible, scaleable procedure to simulate high-dimensional random vectors with arbitrary marginal distributions and a broad class of admissible Spearman correlation and Kendall $\tau$ matrices. We describe the functions included in the package, including multi-core and graphical-processing-unit accelerated algorithms to simulate random vectors, estimate correlation, and compute the nearest correlation matrix. Finally, we demonstrate the power of `bigsimr` by applying these functions to our motivating dataset --- RNA-sequencing data obtained from breast cancer tumor samples with sample size $n=1212$ patients and dimension $d =1026$.
---

>>>>

--- 
title: "Simulating High-Dimensional Multivariate Data"
subtitle: "using the bigsimr R Package"
author:
  - Alfred G. Schissler
  - Edward J. Bedrick
  - Alexander D. Knudson
  - Tomasz J. Kozubowski
  - Tin Nguyen
  - Anna K. Panorska
  - Juli Petereit
  - Walter W. Piegorsch
  - Duc Tran
site: bookdown::bookdown_site
bibliography: ["bigsimr.bib", "packages.bib"]
abstract:  It is critical to realistically simulate data when conducting Monte Carlo studies and evaluating quantitative methods. But measurements are often correlated and high dimensional in this era of big data, however, such as data obtained through high-throughput biomedical experiments. Due to computational complexity and a lack of user-friendly software available to simulate these massive multivariate constructions, researchers may resort to simulation designs that posit independence or perform arbitrary data transformations. This article introduces the `bigsimr` R package that provides a flexible, scaleable procedure to simulate high-dimensional random vectors with arbitrary marginal distributions and a broad class of admissible Spearman correlation and Kendall $\tau$ matrices. We describe the functions included in the package, including multi-core and graphical-processing-unit accelerated algorithms to simulate random vectors, estimate correlation, and compute the nearest correlation matrix. Finally, we demonstrate the power of `bigsimr` by applying these functions to our motivating dataset --- RNA-sequencing data obtained from breast cancer tumor samples with sample size $n=1212$ patients and dimension $d=1026$.
---
